
---

## Kaggle Competition - Predicting prices for Airbnb rentals in NYC

The goal of this competition is to predict the price for airbnb rentals using over 90 variables on the property, host, and past reviews.

## load libraries
```{r}
library(ggplot2)
library(corrplot)
library(randomForest)
library(tidyr)
library(plyr)
library(dplyr)
library(caret)
library(gbm)
library(stringr)
library(xgboost)
```

## Set working directory and input analysisData.csv and scoringData.csv data files

```{r}
setwd("/Users/pratykisallo/Kaggle/AirBNB")
Analysis_Orig <- read.csv(file="analysisData.csv",head=TRUE,sep=",")
Scoring_Orig <- read.csv(file="scoringData.csv",head=TRUE,sep=",")
```


## Data Size and Structure

```{r}
dim(Analysis_Orig)
dim(Scoring_Orig)
```

```{r}
str(Analysis_Orig)
str(Scoring_Orig)
```

## Analysing Response Variable: price
```{r}
ggplot(data = Analysis_Orig[!is.na(Analysis_Orig$price),]) +
geom_histogram(mapping = aes(x = price), binwidth = 10, fill="blue")
summary(Analysis_Orig$price)
```

## Most important numeric variables
```{r}
numericVars <- which(sapply(Analysis_Orig, is.numeric)) #index vector numeric variables
numericVarNames <- names(numericVars) #saving names vector for use later on
cat('There are', length(numericVars), 'numeric variables')

all_numVar <- Analysis_Orig[, numericVars]
cor_numVar <- cor(all_numVar, use="pairwise.complete.obs") #correlations of all numeric variables

cor_sorted <- as.matrix(sort(cor_numVar[,'price'], decreasing = TRUE)) #sort on decreasing correlations with SalePrice

#select only high corelations
CorHigh <- names(which(apply(cor_sorted, 1, function(x) abs(x)>0.3)))
cor_numVar <- cor_numVar[CorHigh, CorHigh]
corrplot.mixed(cor_numVar, tl.col="black", tl.pos = "lt")
```


## Analysis of text variables
```{r}
# these are the important text variables

# summary
# description
# transit
# access
# interaction
# notes
# host_about
# house_rules
# amenities

# online-utility.org was used to do text mining on summary and description
# 'in the heart of' was a top phrase in summary
Analysis_Orig$heart <- grepl("in the heart of", Analysis_Orig$summary) 
Analysis_Orig$heart <- as.numeric(Analysis_Orig$heart)
sum(Analysis_Orig$heart)*100/nrow(Analysis_Orig)
cor(Analysis_Orig$price, Analysis_Orig$heart)

# 'beautiful and spacious' was a top phrase in summary
Analysis_Orig$beautiful <- grepl("beautiful and spacious", Analysis_Orig$summary) 
Analysis_Orig$beautiful <- as.numeric(Analysis_Orig$beautiful)
sum(Analysis_Orig$beautiful)*100/nrow(Analysis_Orig)
cor(Analysis_Orig$price, Analysis_Orig$beautiful)

# 'solo adventurers', 'couples', 'business travelers' were top phrases in description
Analysis_Orig$solo <- grepl("solo adventurers", Analysis_Orig$description) 
Analysis_Orig$solo <- as.numeric(Analysis_Orig$solo)
sum(Analysis_Orig$solo)*100/nrow(Analysis_Orig)
cor(Analysis_Orig$price, Analysis_Orig$solo)

Analysis_Orig$couples <- grepl("couples|romantic", Analysis_Orig$description) 
Analysis_Orig$couples <- as.numeric(Analysis_Orig$couples)
sum(Analysis_Orig$couples)*100/nrow(Analysis_Orig)
cor(Analysis_Orig$price, Analysis_Orig$couples)

Analysis_Orig$business <- grepl("business travelers|business traveler|business", Analysis_Orig$description) 
Analysis_Orig$business <- as.numeric(Analysis_Orig$business)
sum(Analysis_Orig$business)*100/nrow(Analysis_Orig)
cor(Analysis_Orig$price, Analysis_Orig$business)

```

## Neighborhood Variables
```{r}
#neighbourhood, host_location, host_neighborhood, street, state, city, zipcode, market, smartlocation, country, is_location_exact will not be used due to either unclean data or due to imbalanced data

# head(Analysis_Orig$neighbourhood)
# head(Analysis_Orig$host_location)
# head(Analysis_Orig$host_neighborhood)
# head(Analysis_Orig$street)
# head(Analysis_Orig$state)
# head(Analysis_Orig$city)
# head(Analysis_Orig$zipcode)
# head(Analysis_Orig$market)
# head(Analysis_Orig$smartlocation)
# head(Analysis_Orig$country)
# head(Analysis_Orig$is_location_exact)

# neighbourhood_group_cleansed, longitude, latitude are important variables
as.data.frame(table(Analysis_Orig$neighbourhood_group_cleansed))

```


## Missing Values
```{r}
NAcol_Analysis <- which(colSums(is.na(Analysis_Orig)) > 0)
sort(colSums(sapply(Analysis_Orig[NAcol_Analysis], is.na)), decreasing = TRUE)

cat('There are', length(NAcol_Analysis), 'columns with missing values in Analysis set')

NAcol_Scoring <- which(colSums(is.na(Scoring_Orig)) > 0)
sort(colSums(sapply(Scoring_Orig[NAcol_Scoring], is.na)), decreasing = TRUE)

cat('There are', length(NAcol_Scoring), 'columns with missing values in Scoring set')


```

## Splitting the Analysis file

```{r}
set.seed(100) # set seed to ensure results can be replicated
split = sample(1:nrow(Analysis_Orig),0.7*nrow(Analysis_Orig))
train = Analysis_Orig[split,]
test = Analysis_Orig[-split,]
```



## Imputing Missing Values

```{r}
# Cleaning fee imputed with 0
train$cleaning_fee[is.na(train$cleaning_fee)] <- 0
test$cleaning_fee[is.na(test$cleaning_fee)] <- 0
Scoring_Orig$cleaning_fee[is.na(Scoring_Orig$cleaning_fee)] <- 0

# beds imputed with median
train$beds[is.na(train$beds)] <- median(train$beds,na.rm=TRUE)
test$beds[is.na(test$beds)] <- median(test$beds, na.rm=TRUE)
Scoring_Orig$beds[is.na(Scoring_Orig$beds)] <- median(Scoring_Orig$beds, na.rm = TRUE)

# Security deposit imputing with 0 
train$security_deposit[is.na(train$security_deposit)] <- 0
test$security_deposit[is.na(test$security_deposit)] <- 0
Scoring_Orig$security_deposit[is.na(Scoring_Orig$security_deposit)] <- 0

# reviews per month and last_Review
Scoring_Orig$reviews_per_month[is.na(Scoring_Orig$reviews_per_month)] <- 0

Scoring_Orig$last_review <- as.Date(Scoring_Orig$last_review)
Scoring_Orig$last_review[is.na(Scoring_Orig$last_review)] <- "1999-01-01"


```


## Transformation of existing variables

```{r}
# Factor variables to numeric

train$host_is_superhost <- as.numeric(train$host_is_superhost)
test$host_is_superhost <- as.numeric(test$host_is_superhost)
Scoring_Orig$host_is_superhost <- as.numeric(Scoring_Orig$host_is_superhost)

train$instant_bookable <- as.numeric(train$instant_bookable)
test$instant_bookable <- as.numeric(test$instant_bookable)
Scoring_Orig$instant_bookable <- as.numeric(Scoring_Orig$instant_bookable)

train$host_identity_verified <- as.numeric(train$host_identity_verified)
test$host_identity_verified <- as.numeric(test$host_identity_verified)
Scoring_Orig$host_identity_verified <- as.numeric(Scoring_Orig$host_identity_verified)

train$is_business_travel_ready <- as.numeric(train$is_business_travel_ready)
test$is_business_travel_ready<- as.numeric(test$is_business_travel_ready)
Scoring_Orig$is_business_travel_ready <- as.numeric(Scoring_Orig$is_business_travel_ready)

```


```{r}
# Host_since changed to number of days from today

train$host_since <- as.Date(train$host_since)
test$host_since <- as.Date(test$host_since)
Scoring_Orig$host_since <- as.Date(Scoring_Orig$host_since)

train$host_since_days <- difftime(Sys.Date(), train$host_since, units="days")
test$host_since_days <- difftime(Sys.Date(), test$host_since, units="days")
Scoring_Orig$host_since_days <- difftime(Sys.Date(), Scoring_Orig$host_since, units="days")

train$host_since_days <- as.numeric(train$host_since_days)
test$host_since_days <- as.numeric(test$host_since_days)
Scoring_Orig$host_since_days <- as.numeric(Scoring_Orig$host_since_days)


# Calendar_last_scraped changed to number of days from today
train$calendar_last_scraped<- as.Date(train$calendar_last_scraped)
test$calendar_last_scraped<- as.Date(test$calendar_last_scraped)
Scoring_Orig$calendar_last_scraped<- as.Date(Scoring_Orig$calendar_last_scraped)


train$Cal_last_scraped_days <- difftime(Sys.Date(), train$calendar_last_scraped, units="days")
test$Cal_last_scraped_days <- difftime(Sys.Date(), test$calendar_last_scraped, units="days")
Scoring_Orig$Cal_last_scraped_days <- difftime(Sys.Date(), Scoring_Orig$calendar_last_scraped, units="days")

train$Cal_last_scraped_days  <- as.numeric(train$Cal_last_scraped_days)
test$Cal_last_scraped_days  <- as.numeric(test$Cal_last_scraped_days)
Scoring_Orig$Cal_last_scraped_days  <- as.numeric(Scoring_Orig$Cal_last_scraped_days)


# last_review changed to number of days from today
train$last_review<- as.Date(train$last_review)
test$last_review<- as.Date(test$last_review)
Scoring_Orig$last_review<- as.Date(Scoring_Orig$last_review)

train$last_review_days <- difftime(Sys.Date(), train$last_review, units="days")
test$last_review_days <- difftime(Sys.Date(), test$last_review, units="days")
Scoring_Orig$last_review_days <- difftime(Sys.Date(), Scoring_Orig$last_review, units="days")

train$last_review_days  <- as.numeric(train$last_review_days)
test$last_review_days  <- as.numeric(test$last_review_days)
Scoring_Orig$last_review_days  <- as.numeric(Scoring_Orig$last_review_days)


```

```{r}
# Renamed levels in host_response_time
levels(train$host_response_time)[levels(train$host_response_time)=="N/A"] <- "No Reply"
levels(test$host_response_time)[levels(test$host_response_time)=="N/A"] <- "No Reply"
levels(Scoring_Orig$host_response_time)[levels(Scoring_Orig$host_response_time)=="N/A"] <- "No Reply"


```


```{r}
# host_Response_Rate changed to decimal formats
train$host_response_rate<- as.numeric(gsub("%", "", as.character(train$host_response_rate)))/100
train$host_response_rate[is.na(train$host_response_rate)] <-0

test$host_response_rate<- as.numeric(gsub("%", "", as.character(test$host_response_rate)))/100
test$host_response_rate[is.na(test$host_response_rate)] <-0

Scoring_Orig$host_response_rate<- as.numeric(gsub("%", "", as.character(Scoring_Orig$host_response_rate)))/100
Scoring_Orig$host_response_rate[is.na(Scoring_Orig$host_response_rate)] <-0

```

## Creation of new variables


```{r}
# property_type (new variable)
train$property_type_new[train$property_type=="Apartment"]="Apartment" 
train$property_type_new[train$property_type=="Condominium"]="Condominium"
train$property_type_new[train$property_type=="House"]="House"
train$property_type_new[train$property_type=="Loft"]="Loft"
train$property_type_new[train$property_type=="Townhouse"]="Townhouse"
train$property_type_new[which(train$property_type!="Townhouse" & train$property_type!="Apartment" &train$property_type!="Condominium" & train$property_type!="House" & train$property_type!="Loft")]="Others"

test$property_type_new[test$property_type=="Apartment"]="Apartment" 
test$property_type_new[test$property_type=="Condominium"]="Condominium"
test$property_type_new[test$property_type=="House"]="House"
test$property_type_new[test$property_type=="Loft"]="Loft"
test$property_type_new[test$property_type=="Townhouse"]="Townhouse"
test$property_type_new[which(test$property_type!="Townhouse" & test$property_type!="Apartment" & test$property_type!="Condominium" & test$property_type!="House" & test$property_type!="Loft")]="Others"

Scoring_Orig$property_type_new[Scoring_Orig$property_type=="Apartment"]="Apartment" 
Scoring_Orig$property_type_new[Scoring_Orig$property_type=="Condominium"]="Condominium"
Scoring_Orig$property_type_new[Scoring_Orig$property_type=="House"]="House"
Scoring_Orig$property_type_new[Scoring_Orig$property_type=="Loft"]="Loft"
Scoring_Orig$property_type_new[Scoring_Orig$property_type=="Townhouse"]="Townhouse"
Scoring_Orig$property_type_new[which(Scoring_Orig$property_type!="Townhouse" & Scoring_Orig$property_type!="Apartment" & Scoring_Orig$property_type!="Condominium" & Scoring_Orig$property_type!="House" & Scoring_Orig$property_type!="Loft")]="Others"

train$property_type_new <- as.factor(train$property_type_new)
test$property_type_new <- as.factor(test$property_type_new)
Scoring_Orig$property_type_new <- as.factor(Scoring_Orig$property_type_new)

#------- wifi variable

train$wifi<- grepl("Wifi|wifi|internet|Internet", train$amenities)
train$wifi <- as.numeric(train$wifi)

sum(train$wifi)*100/nrow(train)
cor(train$price,train$wifi)

#----- TV variable 
train$TV<- grepl("TV|Cable TV|tv|Cable tv", train$amenities)
train$TV <- as.numeric(train$TV)

test$TV<- grepl("TV|Cable TV|tv|Cable tv", test$amenities)
test$TV <- as.numeric(test$TV)

Scoring_Orig$TV<- grepl("TV|Cable TV|tv|Cable tv", Scoring_Orig$amenities)
Scoring_Orig$TV <- as.numeric(Scoring_Orig$TV)

#-------GYM variable 
train$Gym<- grepl("Gym|gym", train$amenities)
train$Gym <- as.numeric(train$Gym)

sum(train$Gym)*100/nrow(train)
cor(train$price,train$Gym)

test$Gym<- grepl("Gym|gym", test$amenities)
test$Gym <- as.numeric(test$Gym)

Scoring_Orig$Gym<- grepl("Gym|gym", Scoring_Orig$amenities)
Scoring_Orig$Gym <- as.numeric(Scoring_Orig$Gym)


#--------Free Parking
train$parking<- grepl("Free Parking|Free parking|Parking|parking|Free Street Parking|Free street parking", train$amenities)
train$parking <- as.numeric(train$parking)

test$parking<- grepl("Free Parking|Free parking|Parking|parking", test$amenities)
test$parking <- as.numeric(test$parking)

Scoring_Orig$parking<- grepl("Free Parking|Free parking|Parking|parking", Scoring_Orig$amenities)
Scoring_Orig$parking <- as.numeric(Scoring_Orig$parking)

#--------- Smoke detector
train$smoke <- grepl("smoke detector|Smoke Detector|Smoke detector", train$amenities)
train$smoke <- as.numeric(train$smoke)

test$smoke <- grepl("smoke detector|Smoke Detector|Smoke detector", test$amenities)
test$smoke <- as.numeric(test$smoke)

Scoring_Orig$smoke <- grepl("smoke detector|Smoke Detector|Smoke detector", Scoring_Orig$amenities)
Scoring_Orig$smoke <- as.numeric(Scoring_Orig$smoke)

# ------- Family/kids friendly
train$Family <- grepl("Family/kid friendly", train$amenities)
train$Family <- as.numeric(train$Family)

test$Family <- grepl("Family/kid friendly", test$amenities)
test$Family <- as.numeric(test$Family)

Scoring_Orig$Family <- grepl("Family/kid friendly", Scoring_Orig$amenities)
Scoring_Orig$Family <- as.numeric(Scoring_Orig$Family)


# ------- subway ----- 

train$subway <- grepl("Subway|subway|Subways|subways", train$transit)
train$subway <- as.numeric(train$subway)

test$subway <- grepl("Subway|subway|Subways|subways", test$transit)
test$subway <- as.numeric(test$subway)

Scoring_Orig$subway <- grepl("Subway|subway|Subways|subways", Scoring_Orig$transit)
Scoring_Orig$subway <- as.numeric(Scoring_Orig$subway)

# ------- count of amenities
amenities.string <- paste ( train$amenities , collapse = "," )
amenities.vector <- strsplit (amenities.string , "," )[[1]]
amenities.vector.clean <- gsub ( " " , "" , amenities.vector)

#table(amenities.vector.clean)

train$amenities <- as.character(train$amenities)
train$amenities_count <- str_count(train$amenities, ',')
head(train$amenities_count)



amenities.string <- paste ( test$amenities , collapse = "," )
amenities.vector <- strsplit (amenities.string , "," )[[1]]
amenities.vector.clean <- gsub ( " " , "" , amenities.vector)

# table(amenities.vector.clean)

test$amenities <- as.character(test$amenities)
test$amenities_count <- str_count(test$amenities, ',')
head(test$amenities_count)

amenities.string <- paste ( Scoring_Orig$amenities , collapse = "," )
amenities.vector <- strsplit (amenities.string , "," )[[1]]
amenities.vector.clean <- gsub ( " " , "" , amenities.vector)

# table(amenities.vector.clean)

Scoring_Orig$amenities <- as.character(Scoring_Orig$amenities)
Scoring_Orig$amenities_count <- str_count(Scoring_Orig$amenities, ',')
head(Scoring_Orig$amenities_count)


```

## Lasso for feature selection

```{r}
# deleting thumbial, medium, picture, license, monthly, square_feet, weekly_price
train <- select(train, -thumbnail_url, -medium_url, -xl_picture_url, -license, -monthly_price, -square_feet, -weekly_price)
test  <- select(test, -thumbnail_url, -medium_url, -xl_picture_url, -license, -monthly_price, -square_feet, -weekly_price)


Scoring_Orig  <- select(Scoring_Orig, -thumbnail_url, -medium_url, -xl_picture_url, -license, -monthly_price, -square_feet, -weekly_price)

which( colnames(train)=="price" )

# lasso numeric variables for train
numericVars <- which(sapply(train, is.numeric)) #index vector numeric variables
numericVarNames <- names(numericVars) #saving names vector for use later on
train_all_numVar <- train[, numericVars]

which( colnames(train_all_numVar)=="price" )

# lasso numeric variables for test
numericVars <- which(sapply(test, is.numeric)) #index vector numeric variables
numericVarNames <- names(numericVars) #saving names vector for use later on
test_all_numVar <- test[, numericVars]


# lasso numeric variables for Scoring
numericVars <- which(sapply(Scoring_Orig, is.numeric)) #index vector numeric variables
numericVarNames <- names(numericVars) #saving names vector for use later on
Scoring_orig_all_numVar <- Scoring_Orig[, numericVars]

library(glmnet)

x = model.matrix(price~.-1,data=train_all_numVar)
y = train_all_numVar$price

lassoModel = glmnet(x,y, alpha=1)

cv.lasso = cv.glmnet(x,y,alpha=1)

coef(cv.lasso)
```

## Gradient Boosting

```{r}
set.seed(100)
boost = gbm(price~ host_since_days + host_response_time + host_response_rate + host_is_superhost + host_listings_count + host_identity_verified + neighbourhood_group_cleansed + latitude + longitude + property_type_new + room_type + accommodates + bathrooms + bedrooms + beds + bed_type + security_deposit + cleaning_fee + guests_included + extra_people + minimum_nights + maximum_nights + availability_30 + availability_60 + availability_90 + availability_365 + Cal_last_scraped_days + number_of_reviews + last_review_days + review_scores_rating + review_scores_accuracy + review_scores_cleanliness + review_scores_checkin + review_scores_communication + review_scores_location + review_scores_value + instant_bookable + is_business_travel_ready + cancellation_policy + reviews_per_month + amenities_count,data=train,distribution="gaussian",
            n.trees = 100000,interaction.depth = 3,shrinkage = 0.001)

pred_in_sample=predict(boost, newdata = train, n.trees = 100000)
rmse25_in_sample=  sqrt(mean((pred_in_sample-train$price)^2))
rmse25_in_sample
pred_out_of_sample = predict(boost,newdata=test, n.trees = 100000)
rmse25_out_of_sample = sqrt(mean((pred_out_of_sample-test$price)^2)) 
rmse25_out_of_sample 

pred = predict(boost,newdata=Scoring_Orig, n.trees = 100000 )
submissionFile = data.frame(id = Scoring_Orig$id, price = pred)


write.csv(submissionFile, 'sample_submission30.csv',row.names = F)

summary(boost)

```




